---
title: "Untitled"
output: pdf_document
---
```{r}
sm=read.table("student-mat.csv",sep=";",header=TRUE)
library(ggplot2)
```

IV. Regression

In this report, we would like to find factors that are related to students academic performance. We will perform a multivariate regression analysis and use step-wise variable selection to find significant variables related to our response variable (academic performance - "Score"). Prior to the regression analysis, we want to get a general idea about the relationship between the variables. There are several things we want to inspect :
1) Distribution of response variable
if there is skewness in the distribution, we should consider transforming the variable (i.e logarithmic)
2) Correlation between response variable and each explanatory variables
3) Correlation among explanatory variables
if there is a high correlation among explanatory variables, then there is possible multicollinearity problem

After performing the inspection and do the necessary action, we will continue with the regression and check the validity of the model.


I. Distribution of Response Variable
```{r setup, include=FALSE}
ggplot(sm) + geom_density(aes(x=Score)) + labs(x = "Academic Score", y="% Frequency", title = "Academic Score of Secondary School Student")
```
From the graph, we can see that the distribution is slightly left-skewed. Generally, we would transform the variable to become more normally distributed. However, for the purpose of this project, we will not do any transfomation on the response variable and assume that the distribution is normal.

II. Correlation between response variable and each explanatory variables
```{r setup, include=FALSE}
panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor, ...)
{
  usr <- par("usr"); on.exit(par(usr))
  par(usr = c(0, 1, 0, 1))
  r <- abs(cor(x, y))
  txt <- format(c(r, 0.123456789), digits = digits)[1]
  txt <- paste0(prefix, txt)
  if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
  text(0.5, 0.5, txt, cex = cex.cor * r)
}

pairs(sm[,c("G3", "G1","G2","school","sex","age", "address","famsize","Pstatus","Medu","Fedu",
            "Mjob","Fjob","reason","guardian","traveltime","studytime","failures",
            "schoolsup","famsup","paid","activities","nursery","higher","internet",
            "romantic","famrel","freetime","goout","Dalc","Walc","health","absences")],
      lower.panel = panel.smooth, upper.panel = panel.cor)
```

(remi's part)

III. Correlation between response variable and each explanatory variables

```{r setup, include=FALSE}
panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor, ...)
{
  usr <- par("usr"); on.exit(par(usr))
  par(usr = c(0, 1, 0, 1))
  r <- abs(cor(x, y))
  txt <- format(c(r, 0.123456789), digits = digits)[1]
  txt <- paste0(prefix, txt)
  if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
  text(0.5, 0.5, txt, cex = cex.cor * r)
}

pairs(sm[,c("G3", "G1","G2","school","sex","age", "address","famsize","Pstatus","Medu","Fedu",
            "Mjob","Fjob","reason","guardian","traveltime","studytime","failures",
            "schoolsup","famsup","paid","activities","nursery","higher","internet",
            "romantic","famrel","freetime","goout","Dalc","Walc","health","absences")],
      lower.panel = panel.smooth, upper.panel = panel.cor)
```

(remi's part)


```{r setup, include=FALSE}

#2. derive variables

# father or mother job is teacher = 1
sm$f_or_mteach <- 0
sm[j,"f_or_mteach"] <- as.character(sm[,"f_or_mteach"])
for (j in 1:nrow(sm)) {

  if (sm[j, "Fjob"] == "teacher") sm[j,"f_or_mteach"] <- 1
  if (sm[j, "Mjob"] == "teacher") sm[j,"f_or_mteach"] <- 1
}

sm[,"f_or_mteach"] 

pairs(sm[,c("G3","f_or_mteach")], lower.panel = panel.smooth, upper.panel = panel.cor)


# father or mother earns university degree = 1 
sm$f_or_medu4 <- 0
sm[j,"f_or_medu4"] <- as.character(sm[,"f_or_medu4"])
for (j in 1:nrow(sm)) {

  if (sm[j, "Medu"] == 4) sm[j,"f_or_medu4"] <- 1
  if (sm[j, "Fedu"] == 4) sm[j,"f_or_medu4"] <- 1
}

sm[,"f_or_medu4"] 
pairs(sm[,c("G3","f_or_medu4")], lower.panel = panel.smooth, upper.panel = panel.cor)

```


REGRESSION
```{}

# academic performance is measured by the average of exam scores
sm[, "Score"] <- (sm[, "G1"] + sm[, "G2"] + sm[, "G3"])/3

#Initial regression

reg <- lm(Score ~ school + sex + age + address + famsize + Pstatus + Medu + Fedu + 
                Mjob + Fjob + reason + guardian + traveltime + studytime + failures + 
                schoolsup + famsup + paid + activities + nursery + higher + internet + 
                romantic + famrel + freetime + goout + Dalc + Walc + health + absences +
                f_or_mteach + f_or_medu4
                , data = sm)
summary(reg)

# Step wise variable selection
reg1 <- update(reg, . ~ . - guardian)
summary(reg1)
reg2 <- update(reg1, . ~ . - Fjob)
summary(reg2)
reg3 <- update(reg2, . ~ . - Medu)
summary(reg3)
reg4 <- update(reg3, . ~ . - Fedu)
summary(reg4)
reg5 <- update(reg4, . ~ . - famrel)
summary(reg5)
reg6 <- update(reg5, . ~ . - Walc)
summary(reg6)
reg7 <- update(reg6, . ~ . - Dalc)
summary(reg7)
reg8 <- update(reg7, . ~ . - paid)
summary(reg8)
reg9 <- update(reg8, . ~ . - nursery)
summary(reg9)
reg10 <- update(reg9, . ~ . - Pstatus)
summary(reg10)
reg11 <- update(reg10, . ~ . - school)
summary(reg11)
reg12 <- update(reg11, . ~ . - activities)
summary(reg12)
reg13 <- update(reg12, . ~ . - f_or_mteach)
summary(reg13)
reg14 <- update(reg13, . ~ . - internet)
summary(reg14)
reg15 <- update(reg14, . ~ . - Mjob)
summary(reg15)
reg16 <- update(reg15, . ~ . - reason)
summary(reg16)
reg17 <- update(reg16, . ~ . - age)
summary(reg17)
reg18 <- update(reg17, . ~ . - traveltime)
summary(reg18)
reg19 <- update(reg18, . ~ . - freetime)
summary(reg19)
reg20 <- update(reg19, . ~ . - famsize)
summary(reg20)
reg21 <- update(reg20, . ~ . - absences)
summary(reg21)
reg22 <- update(reg21, . ~ . - health)
summary(reg22)
reg23 <- update(reg22, . ~ . - higher)
summary(reg23)
reg24 <- update(reg23, . ~ . - romantic)
summary(reg24)


#To cross check on the final model and the significant variables, we use existing R library:

library(MASS)
step <- stepAIC(reg, direction="both")

#Final Model

library(stargazer)
stargazer(reg24, type = "text", digits = 2)

```

```{}

# residual plot
fit_res <-  data.frame(sm, fitted.value = fitted(reg24), residual = resid(reg24))

ggplot(data = fit_res) + geom_point(aes(x = Score, y = residual)) + geom_hline(yintercept = 0) + labs(x = "Academic Score", y= "residual", title = "Residual Plot of Academic Score Regression")
```


from the regression result, we can see that alcohol consumption is not one of the significant variables that explains student academic performance. Instead, they are:
1) sex
2) address
3) study time
4) number of past class failure
5) extra educational school support
6) family educational support
7) frequency of going out with friends
8) parents level of education

However, statistically, the model itself does not explain academic score very well, as we see through the R2 result as well as the residual plot graph. the r2 is 21.74%, which means that the model can only explain 21.74% of the variance in student's academic score. while the residual plot graph does not show a random scattered plot.

there are other significant factors not available in the dataset that we think can explain score better. for example, IQ and EQ score, nutrition, or time spent for physical exercise.

