---
title: "Students' Alcohol Consumption impacts on academic Results on the example of a secondary Scool Maths Class in Portugal"
output: github_document
Data set source: P. Cortez and A. Silva. Using Data Mining to Predict Secondary School Student Performance. In A. Brito and J. Teixeira Eds., Proceedings of 5th FUture BUsiness TEChnology Conference (FUBUTEC 2008) pp. 5-12, Porto, Portugal, April, 2008, EUROSIS, ISBN 978-9077381-39-7. 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Students' Alcohol Consumption impacts on academic Results on the example of a secondary Scool Maths Class in Portugal

## Introduction 

After the last financial crisis, the European Union has recognised that investments are required in order to reverse the latest trends in economic growth decline by creating new jobs and a better educated workforce. Hence, the European Comission has set up a ‘Strategy for smart, sustainable and inclusive growth’ (European Comission, 2010), where a crucial part of investment is to contribute to the education of the next generations: The aim is to reduce the share of early school leavers to below 10% and ensure that at least 40% of the younger generation obtain a tertiary degree. In the following, the current educational condition in Europe will be investigated and an examplary deeper analysis on one of the greater affected countries, Portugal, will be provided in order to gain an understanding of factors influencing the academic achievements in secondary school, which are crucial for students to enter tertiary education and subsequently performing well in their working career. 


## Description of Dataset

```{r pressure, echo=FALSE}
sm <- read.csv("student-mat.csv", sep=';')
```


## Descriptive Statistics 

You can include R code in the document as follows:

```{r cars}
summary(cars)
```

## Correlation Analysis

Text

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.


